Y---
title: "DS Project EDA (mental health)"
author: "Chaelin Shin"
date: "March 2020"
output: html_document
---


```{r setup, include=T}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
# knitr::opts_chunk$set(warning = F, results = 'markup', message = F)
knitr::opts_chunk$set(warning = F, results = 'hide', message = F)
# knitr::opts_chunk$set(include = F)
# knitr::opts_chunk$set(echo = TRUE)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

```{r basic, include=T}
# use this function to conveniently load libraries and work smoothly with knitting
# can add quietly=T option to the require() function
loadPkg = function(pkg, character.only = FALSE) { 
  if (!character.only) { pkg <- as.character(substitute(pkg)) }
  pkg <- ifelse(!character.only, as.character(substitute(pkg)) , pkg)  
  if (!require(pkg,character.only=T, quietly =T)) {  install.packages(substitute(pkg),dep=T); if(!require(pkg,character.only=T)) stop("Package not found") } 
}
loadPkg(knitr)

# unload/detact package when done using it
unloadPkg = function(pkg, character.only = FALSE) { 
  if(!character.only) { pkg <- as.character(substitute(pkg)) } 
  search_item <- paste("package", pkg,sep = ":") 
  while(search_item %in% search()) { detach(search_item, unload = TRUE, character.only = TRUE) } 
}
```

``` {r import_data, include=T}
# reading data

getwd()
health <- read.csv("samadult.csv",header=TRUE)

```

```{r outlierKD, include=T}
# oulier function in case there are any outliers to be removed
outlierKD <- function(dt, var) { 
     var_name <- eval(substitute(var),eval(dt))
     na1 <- sum(is.na(var_name))
     m1 <- mean(var_name, na.rm = T)
     par(mfrow=c(2, 2), oma=c(0,0,3,0))
     boxplot(var_name, main="With outliers")
     hist(var_name, main="With outliers", xlab=NA, ylab=NA)
     outlier <- boxplot.stats(var_name)$out
     mo <- mean(outlier)
     var_name <- ifelse(var_name %in% outlier, NA, var_name)
     boxplot(var_name, main="Without outliers")
     hist(var_name, main="Without outliers", xlab=NA, ylab=NA)
     title("Outlier Check", outer=TRUE)
     na2 <- sum(is.na(var_name))
     cat("Outliers identified:", na2 - na1, "n")
     cat("Propotion (%) of outliers:", round((na2 - na1) / sum(!is.na(var_name))*100, 1), "n")
     cat("Mean of the outliers:", round(mo, 2), "n")
     m2 <- mean(var_name, na.rm = T)
     cat("Mean without removing outliers:", round(m1, 2), "n")
     cat("Mean if we remove outliers:", round(m2, 2), "n")
     response <- readline(prompt="Do you want to remove outliers and to replace with NA? [yes/no]: ")
     if(response == "y" | response == "yes"){
          dt[as.character(substitute(var))] <- invisible(var_name)
          assign(as.character(as.list(match.call())$dt), dt, envir = .GlobalEnv)
          cat("Outliers successfully removed", "n")
          return(invisible(dt))
     } else{
          cat("Nothing changed", "n")
          return(invisible(var_name))
     }
}
```

``` {r subset_occpn, include=F}
# subset by occupation (6414 observations)
# six occupation: business & finance, computer & math, life physical & social science, education, art & design, production
# change occupation variables to factors
# change name of levels

health6 <- subset(health, health$OCCUPN2 %in% c(2,3,5,8,9,21))
health6$OCCUPN2 <- as.factor(health6$OCCUPN2)

levels(health6$OCCUPN2) <- c("Bus","Comp","Science","Educ","Art","Prod")

```

# EDA for Intro to Data Science Project I: What is the healthiest occupation in the U.S.?

``` {r subset, include=T}
# how many people are there within each subset? 
# variables: 
#  - ALDURA17: duration (in years) of depression/anxiety/emotional problem
#  - ALDURA29: how long have you had an alcohol/drug/substance abuse problem? 
#  - ASINERV: during the past 30 days, how often did you feel nervous? 
#  - ASIHOPLS: during the past 30 days, how often did you feel hopeless?
#  - ASIWTHLS: during the past 30 days, how often did you feel worthless? 
# Trying to see how many people within the subset, health6, have mental health problems, substance abuse problems, or have been feeling nervous or hopeless. 
# Eliminated unnecessary outliers for mental health subset (96: unknown number of years, 97: refused, 98: not ascertained, 99: don't know)

MentalHealth <- subset(health6, ALDURA17 > 0 & ALDURA17 < 96)

Substance <- subset(health6, ALDURA29 > 0)

Nervous <- subset(health6, ASINERV > 0 & ASINERV < 7)

Hopeless <- subset(health6, ASIHOPLS > 0 & ASIHOPLS < 7)


# MentalHealth: 132 observations
# Substance: 0 --> excluded from our analysis
# Nervous: 6231
# Hopeless: 6229


```

## Depression/anxiety/emotional problem 

In order to check normality, below are the histogram and qq-plot of the mental health variable (duration of depression/anxiety/emotional problem): 

``` {r depression, include=T}
# summary of the MentalHealth subset
summary(MentalHealth$ALDURA17)

# normality check - histogram and qqplot
loadPkg("ggplot2")

# histogram

ggplot(MentalHealth,aes(x=ALDURA17))+
  geom_histogram(color="magenta",fill='mistyrose')+
  labs(title="Histogram of Mental Health Variable",x="Duration of Mental Health (years)")

# qqplot

qqnorm(MentalHealth$ALDURA17,main="QQ-Plot of Mental Health Variable",col="magenta")
qqline(MentalHealth$ALDURA17)
```


The histogram does not seem to be perfectly normal, but rather, slightly skewed to the right. And the qq-plot does not completely resemble the qq-line, but it is fairly close to it. As long as it's not severely non-normal, we can move on to do the ANOVA test.

Although it does not look like there are outstanding outliers to remove, let's still try removing the outliers, if any. 

``` {r without outlier, include=T}
# remove outliers 
# 0 outliers
outlierKD(MentalHealth, ALDURA17)

# histogram

ggplot(MentalHealth,aes(x=ALDURA17))+
  geom_histogram(color="magenta",fill='mistyrose')+
  labs(title="Histogram of Mental Health Variable",x="Duration of Mental Health (years)")


```


There were no outliers identified and thus, the histogram looks exactly the same as above. 


Below is the boxplot to see the relationship between the occupations and mental 
health problems: 

``` {r boxplot_mental health,include=T}
# boxplot 
ggplot(MentalHealth, aes(x=OCCUPN2,y=ALDURA17)) + geom_boxplot() + geom_boxplot(col=rainbow(6),fill=rainbow(6,alpha=0.2),outlier.color="red",outlier.shape=8, outlier.size=4) + labs(title="Mental Health vs. Occupation",x="OCCUPATION",y="Mental Health")

```


It looks like people with the production occupation have more mental health problems. However, this result is not too reliable considering that there are only 132 people among 6414 who answered they have mental health problems in the survey. 


Let's try running the ANOVA test:
``` {r anova_depression, include=T}
# ANOVA test 

anovaM = aov(ALDURA17~OCCUPN2,data=MentalHealth)
summary(anovaM)

```

The results are not too impressive. The F value is 1.62 and the P-value is 0.16, thus statistically insignificant.


## Feeling Nervous 

Below is the bar chart to see the responses to the question asking how often one felt nervous in the past 30 days:

``` {r nervous, include=T}
# in the past 30 days, how often did you feel nervous? 
# make factor variable
#  -  1: ALL of the time
#  -  2: MOST of the time
#  -  3: SOME of the time
#  -  4: A LITTLE of the time
#  -  5: NONE of the time
# change name of levels 

Nervous$ASINERV <- as.factor(Nervous$ASINERV)
levels(Nervous$ASINERV) <- c("All","Most","Some","Little","None")

# summary of the Nervous subset
summary(Nervous$ASINERV)

# bar chart
ggplot(Nervous,aes(x=ASINERV))+
  geom_bar(color="magenta",fill="mistyrose")+
  labs(x="How often do you feel nervous?")


```


Most of the people answered that they did not feel nervous in the past 30 days. However, there were more than 1000 people saying that they felt a little nervous and a little less than 1000 people saying that they were nervous some of the time.


Let's try running the chi-squared test: 

``` {r chi-squared test_nervous, include=T}

loadPkg("MASS")
loadPkg("corrplot")

#contingency table
contableNerv = table(Nervous$OCCUPN2,Nervous$ASINERV)

chitest = chisq.test(contableNerv)
chitest

chitest$statistic
chitest$parameter
chitest$expected
pchisq(116,20,lower.tail=F)

```


The p-value is 1.56e-15, which tells us that occupation and nervousness are not independent. 

Below is a plot that shows us the relationship between each occupation and nervousness: 

``` {r nervous plot, include=T}
corrplot(chitest$residuals,is.cor=FALSE)

```


The production occupation has the strongest blue circle, which indicates that this occupation had the most positive relationship with nervousness. 


## Feeling Hopeless

Below is the bar chart to see the responses to the question asking how often one felt hopeless in the past 30 days:

``` {r hopeless, include=T}
# make factor variable
#  -  1: ALL of the time
#  -  2: MOST of the time
#  -  3: SOME of the time
#  -  4: A LITTLE of the time
#  -  5: NONE of the time
# change name of levels

Hopeless$ASIHOPLS <- as.factor(Hopeless$ASIHOPLS)
levels(Hopeless$ASIHOPLS) <- c("All","Most","Some","Little","None")

# summary of the Nervous subset
summary(Hopeless$ASIHOPLS)

# bar chart
ggplot(Hopeless,aes(x=ASIHOPLS))+
  geom_bar(color="magenta",fill="mistyrose")+
  labs(x="How often do you feel hopeless?")


```


Most of the people did not feel hopeless in the past 30 days. Overall, there are lot less people who felt hopeless than who felt nervous.


Let's try running the chi-squared test. 

``` {r chi-squared test_hopeless, include=T}

#contingency table
contableHope = table(Hopeless$OCCUPN2,Hopeless$ASIHOPLS)

chitest2 = chisq.test(contableHope)
chitest2

chitest2$statistic
chitest2$parameter
chitest2$expected
pchisq(64.8,20,lower.tail=F)


```


The p-value is 1.26e-06, which tells us that occupation and hopelessness are not independent. 

Below is a plot that shows us the relationship between each occupation and hopelessness: 

``` {r hopeless plot, include=T}
corrplot(chitest2$residuals,is.cor=FALSE)

```


Again, the production occupation has the strongest blue circles, which tells us that they felt the most hopeless. 

## How much do the results differ between male and female (within the six occupations)? 

### Depression/Anxiety/Emotional Problem 

``` {r male_depression, include=T}

mental_M <- subset(MentalHealth, MentalHealth$SEX==1)
mental_F <- subset(MentalHealth, MentalHealth$SEX==2)

# 46 observations for mental_M
# 83 observations for mental_F
```


Below are the boxplots for male and female:

``` {r boxplot_depression_gender, include=T}

# boxplot for male 
ggplot(mental_M, aes(x=OCCUPN2,y=ALDURA17)) + geom_boxplot() + geom_boxplot(col=rainbow(6),fill=rainbow(6,alpha=0.2),outlier.color="red",outlier.shape=8, outlier.size=4) + labs(title="Mental Health vs. Occupation (Male)",x="OCCUPATION",y="Mental Health")

# boxplot for female
ggplot(mental_F, aes(x=OCCUPN2,y=ALDURA17)) + geom_boxplot() + geom_boxplot(col=rainbow(6),fill=rainbow(6,alpha=0.2),outlier.color="red",outlier.shape=8, outlier.size=4) + labs(title="Mental Health vs. Occupation (Female)",x="OCCUPATION",y="Mental Health")

```


### Feeling Nervous


Below is a bar chart that shows how many male and female within the six occupations felt nervous:



``` {r gender_nervous, include=T}

# Make sex variable factors
Nervous$SEX <- as.factor(Nervous$SEX)
# Change name of levels
levels(Nervous$SEX) <- c("Male","Female")

# bar chart

ggplot(Nervous)+
  geom_bar(aes(x=ASINERV,fill=SEX),position = position_dodge(preserve = 'single')) + labs(x="How often do you feel nervous?")


```


### Feeling Hopeless

``` {r gender_hopeless, include=T}

# Make sex variable factors
Hopeless$SEX <- as.factor(Hopeless$SEX)
# Change name of levels
levels(Hopeless$SEX) <- c("Male","Female")

# bar chart

ggplot(Hopeless)+
  geom_bar(aes(x=ASIHOPLS,fill=SEX),position = position_dodge(preserve = 'single')) + labs(x="How often do you feel hopeless?")


```

``` {r unload_package,include=T}
#unload packages

unloadPkg("ggplot2")
unloadPkg("MASS")
unloadPkg("corrplot")

```

